State of the Practice for Medical Imaging Software Based on Open Source Repositories

### 2024-11-08 StateOfPractice/Papers/SOP_MI_CritReviewsBioMedEng/SOP_MI_OSR.tex
to the end of the paper

At least to some extent the problems for MI software users are due to holes in
their background knowledge.  For research software in general, users do not
always have the expertise required to install or use the software
\cite{WieseEtAl2019}. A similar pattern has been observed during interviews with
LBM software developers [citation redacted for blind review].
%\cite{SmithEtAl2024}.

@@

\begin{itemize}
    \item Use documentation (user manuals, mailing lists, forums) (mentioned by
    4 developers)
    \item Usability tests and interviews with end users; and, (mentioned by 3
    developers)
    \item Adjusting the software according to user feedback. (mentioned by 3
    developers)
\end{itemize}

@@

\section{Software Qualities}\label{sec:sq}
In addition to pain points, we summarize
MI developer strategies for improving maintainability and reproducibility.
Although the interviewees did not explicitly identify these two qualities as
pain points, we did discuss threats to these qualities and ways to improve them
as part of our interview process \cite{SmithEtAl2021}.  The interviewee's
practices for addressing pain points and improving quality can potentially be
emulated by other MI developers. Moreover, these practices may provide examples
%HERE
that can be followed by other research software domains.
Up to this point, we have covered the pain points that came up in interviews
with MI developers, along with a summary of the techniques that are currently
used to address these pain points.  Although the developers did not explicitly
identify the qualities of maintainability and reproducibility as pain points in
our interviews, as part of our interview questions
(Section~\ref{sec_interview_methods}) they did share their approaches for
improving these qualities, as discussed below.

\begin{enumerate} \item[Q\refstepcounter{qnum}\theqnum
\label{Q_Maintainability}:] \textbf{Maintainability:}  Maintainability has been
rated as the third most important software quality for research software
\cite{Nguyen-HoanEtAl2010}. The push for sustainable software
\cite{deSouzaEtAl2019} is motivated by the pain that past developers have had
with accumulating too much technical debt \cite{KruchtenEtAl2012}.  For LBM
software, technical debt has been identified as one of the developer pain points
[citation redacted for blind review]. %\cite{SmithEtAl2024}.

To improve maintainability, the most popular (with five out of nine interviewees
mentioning it) strategy is to use a modular approach, with often repeated
functions in a library.  Other strategies that were mentioned for improving
maintainability include supporting third-party extensions, an easy-to-understand
architecture, a dedicated architect, starting from simple solutions, and
documentation.  The \textit{3D Slicer} team used a well-defined structure for
the software, which they named as an ``event-driven MVC pattern''.  \textit{3D
Slicer} discovers and loads necessary modules at runtime. The \textit{BioImage
Suite Web} team had designed and re-designed their software multiple times in
the last 10+ years. They found that their modular approach effectively supports
maintainability \cite{Joshi2011}.

\item[Q\refstepcounter{qnum}\theqnum \label{Q_Reproducibility}:]
\textbf{Reproducibility:}  Although the MI developers did not mention
reproducibility explicitly as a pain point, they did mention the need to improve
documentation.  Good documentation does not just address the pain points of lack
of developer time (\ppref{P_LackDevTime}), technology hurdles
(\ppref{P_TechnologyHurdles}), usability \ppref{P_Usability}, and
maintainability.  Documentation is also necessary for reproducibility. The
challenges of inadequate documentation are a known problem for research software
\cite{PintoEtAl2018, WieseEtAl2019} and for non-research software
\cite{LethbridgeEtAl2003}.

In our interviews, we discussed threats to reproducibility and strategies for
improving it.  The threats that were mentioned include closed-source software,
no user interaction tests, no unit tests, the need to change versions of some
common libraries, variability between CPUs, and misinterpretation of how
manufacturers create medical images.

The most commonly cited (by 6 teams) strategy to improve reproducibility was
testing (regression tests, unit tests, having good tests). The second most
common strategy (mentioned by 5 teams) is making code, data, and documentation
available, possibly by creating open-source libraries.  Other ideas that were
mentioned include running the same tests on all platforms, a dockerized version
of the software to insulate it from the OS environment, using standard
libraries, monitoring the upgrades of the library dependencies, clearly
documenting the version information, bringing along the exact versions of all
the dependencies with the software, providing checksums of the data, and
benchmarking the software against other software that overlaps in functionality.
Specifically one interviewee suggested using \textit{3D Slicer} as the benchmark
to test their reproducibility.

@@

In this section we provide recommendations to address the pain points from
Section~\ref{painpoints} to answer~\rqref{RQ_Recommend}.  Our recommendations
are not lists of criticisms for what should have been done in the past, or what
should be done now; they are suggestions for consideration in the future. We
expand on some of the ideas that came out of our interviews with developers
(Section~\ref{painpoints}), including continuous integration, moving to web
applications, and enriching the test data sets. We also bring in new like
employing linters, peer review, and design for change.  Our aim is to mention
ideas that are at least somewhat beyond conventional best practices. The ideas
listed here have the potential to become best practices in the medium to
long-term. We list the ideas roughly in the order of increasing implementation
effort.

@@

Continuous integration involves frequent pushes to a code repository.  With
every push the software is built and tested \cite[p.\ 13]
{HumbleAndFarley2010}, \cite{ShahinEtAl2017, Fowler2006}.  CI can take
significant time and effort to set up and integrate into a team's workflow, but
the benefits are impressive, as follows:

\begin{itemize}
    \item Elimination of headaches associated with a separate integration phase
    \cite{Fowler2006}, \cite[p.\ 20]{HumbleAndFarley2010}.
    \item Detection and removal of bugs \cite{Fowler2006} via
    automated testing.  
    \item Everyone is always working on a stable base, since the rejection of
    inadequate commits means that the main branch will always be working.  A
    stable base will always pass all tests.  If the CI system uses generators
    and linters, it will also have current documentation and standard compliant
    code.  A stable base improves developer productivity, allowing them to focus
    on coding, testing, and documentation.
\end{itemize}

Continuous integration and delivery helps with addressing several pain points.
For instance, CI/CD helps reduce development time (\ppref{P_LackDevTime}) by
removing the need for a time-consuming integration stage and by automating
regression testing.  Automated regression tests also help with ensuring
correctness (\ppref{P_Correctness}) and the quality of reproducibility
(\qref{Q_Reproducibility}). Details on CI for MI can be found in [citation
removed for blind review] %Smith et al. \cite{SmithEtAl2024_MI_SOP} 
, including an overview of the elements of CI, how to enable incorporation of CI
into a team's workflow, how to set-up a CI system, and challenges for CI.

\subsection{Move To Web Applications} \label{sec_webapps}

Section~\ref{painpoints} describes the pain point of technology hurdles
(\ppref{P_TechnologyHurdles}), which motivates considering the use of web
applications. Here we give further advice to help with deciding whether to adopt
a web application. The decision will be based on whether, on balance, the web
application improves the four factors identified by developers: compatibility,
maintainability, performance, and security. To enable decision-making, a team
will need to prioritize between these factors, based on their objectives and
experience. The suggestions are intended to provide ideas and avenues for
exploration; a web application will not be the right fit for all projects and
all teams.

@@

\item \textbf{Modern technologies may improve frontend performance.} Web
applications with only a frontend usually perform worse than native
applications. However, new technologies may ease this difference. For example,
some JavaScript libraries can help the frontend harness the power of the
computer's GPU and accelerate graphical computing. In addition, there are new
frameworks helping developers with cross-platform compatibility. For example,
the \href{https://flutter.dev/}{Flutter} project enables support for web,
mobile, and desktop OS with one codebase.  Other options include
\href{https://vuejs.org/} {Vue}, \href{https://angular.io/} {Angular} and
\href{https://reactjs.org/} {React}, and \href{https://elm-lang.org/}{Elm}.

\item \textbf{Backend servers can potentially deliver high performance.} Web
applications with backend servers may perform even better than native
applications. If a team needs to support lower-end computers, it is good to use
back-end servers for heavy computing tasks.  For backend servers where traffic
and latency is not an issue, options include
\href{https://www.django-rest-framework.org/} {Django}, \href{https://laravel.com/} {Laravel} and
\href{https://nodejs.org/en/} {Node.js}.  The advantage of Django is that it provides access to Python
libraries.  For backend servers where traffic and latency is an issue,
\href{https://github.com/gin-gonic/gin} {Gin} is an option.

\item \textbf{Backend servers can have low costs.} Serverless solutions from
major cloud service providers (like Amazon Web Services (AWS) and Google Cloud
Platform) may be worth exploring. Serverless solutions still use a server, but
the server provider only charges the team when they use the server. The solution
is event-driven, and costs the team by the number of requests processed. Thus,
serverless can be very cost-effective for less intensively used functions.

\item \textbf{Web transmission may diminish security.} Transferring sensitive
data on-line can be a problem for projects requiring high security. Regulations
for some MI applications may forbid doing web transmissions. In this case, a web
application with a backend may not be an option.

@@

As described in Section~\ref{painpoints}, ensuring correctness
(\ppref{P_Correctness}) via testing can be problematic because of limited access
to real-world medical imaging datasets.  We build on the suggestions we heard
from our interviewees as follows:

\begin{itemize}
\item \textbf{Build and maintain good connections to datasets.} A team can build
connections with professionals working in the medical domain, who may have
access to private datasets and can perform tests for the team.

\item \textbf{Collect and maintain datasets over time.} A team may face problems
caused by various unique inputs over the years of software development. This
data should be collected and maintained over time to form a good, comprehensive,
dataset for testing.

\item \textbf{Search for open data sources.} In general, there are many open MI
datasets.  For instance, there are
\href{https://nihcc.app.box.com/v/ChestXray-NIHCC}{Chest X-ray Datasets} by
National Institute of Health \cite{WangEtAl2017},
\href{https://www.cancerimagingarchive.net/}{Cancer Imaging Archive}
\cite{PriorEtAl2017}, \href{https://medpix.nlm.nih.gov/home}{MedPix} by
National Library of Medicine \cite{Smirniotopoulos2014}, and datasets for liver
\cite{BilicEtAl2019} and brain \cite{MenzeEtAl2015} tumor segmentation
benchmarks.

\item \textbf{Create sample data for testing.} If a team can access tools
creating sample data, they may also self-build datasets for testing. For
example, an MI software development team can use an MRI scanner to create images
of objects, animals, and volunteers. The team can build the images based on
specific testing requirements.

\item \textbf{Remove privacy from sensitive data.} For data with sensitive
information, a team can ask the data owner to remove such information or add
noise to protect privacy. One example is using de-identified copies of medical
images for testing.

@@

\subsection{Employ Linters} \label{Sec_Linters}

A linter is a tool that statically analyzes code to find programming errors,
suspicious constructs, and stylistic inconsistencies \cite{Wikipedia2022_Lint}.
Linters can be used as an ad hoc check for code files, but they really come into
their own when used as part of a CI system, as discussed in
Section~\ref{Sec_ContinuousIntegration}. Almost none of the research software
guidelines that we consulted, summarized in Section~\ref{Sec_CompareArtifacts},
mention linters.  Despite the lack of mention in the guidelines, we believe that
linters have the potential to improve code quality at a relatively low cost.

Linters have the following benefits: finding potential bugs, finding memory
leaks, improving performance, standardizing code with respect to formatting,
removing silly errors before code reviews, and catching potential security
issues \cite{SourceLevel2022_Lint}. Most popular programming languages have an
accompanying linter.  For example, Python has the options of PyLint, flake8 and
Black \cite{Zadka2018}.

Linters address several MI pain points.  For instance, linters address the lack
of development time (\ppref{P_LackDevTime}) by increasing the developer's
productive time via guarding against frustrating, time-consuming, and mundane
mistakes.  Moreover, since a linter can include rules that capture the wisdom of
senior programmers, it can help newer developers avoid common mistakes. With
respect to the technology hurdle pain point (\ppref{P_TechnologyHurdles}),
linters can assist with the move toward web applications
(Section~\ref{sec_webapps}).  For instance, ESLint in React is a pluggable
linter that lets the developer know if they have imported something and not used
it, if a function could be short-handed, if there are indentation
inconsistencies, etc. By insisting on code
standardization, linters can reduce technical debt and thus improve
maintainability (\qref{Q_Maintainability}). Although linters are tools for code
analysis, the idea of statically checking for adherence to basic rules can be
extended to check documentation. Tools to enforce documentation standards
partially explains the relatively higher quality of statistical tools that are
part of the Comprehensive R Archive Network (CRAN) [citation redacted for blind
review].  %\cite{SmithEtAl2018_StatSoft}.

\subsection{Conduct a Mix of Rigorous and Informal Peer Reviews} \label{Sec_PeerReview}

We advocate incorporating peer review into the development process, as
frequently recommended for research software \cite{HerouxEtAl2008,
OrvizEtAl2017, USGS2019}. In most cases a modern, lightweight review, should be
adequate.  Modern code review is informal, tool-based, asynchronous, and focused
on reviewing code changes \cite{SadowskiEtAl2018}. Managing a project via GitHub
pull requests is an example of a modern approach to reviewing code.  Software
development organizations have moved to this lightweight style of code review
because of the inefficiencies of rigorous inspections \cite{RigbyAndBird2013}.
However, for important parts of the code, developers may benefit from mixing in
a more rigorous approach.

Fagan \cite{Fagan1976} began work on rigorous review via code inspection.
Elements of a typical inspection include reviewing the code against a checklist
(checking the consistency of variable names, look for terminating loops, etc.),
performing specific review tasks (such as summarizing the code's purpose,
cross-referencing the code to the technical manual, creating a data dictionary
for a given module, etc.) Rigorous inspection finds 60-65\% of latent defects on
average, and often tops 85\% in defect removal efficiency \cite{Jones2008}. The
success rate of code inspection is generally higher than most forms of testing,
which average between 30 --- 35\% for defect removal efficiency
\cite{EbertAndJones2009, Jones2008}. For research software, Kelly and Shepard
\cite{KellyAndShepard2000} show a task based inspection approach can be
effective.

Due to improving code quality and increasing knowledge transfer, peer review
addresses the same pain points and qualities as linters
(Section~\ref{Sec_Linters}): \ppref{P_LackDevTime}, \ppref{P_TechnologyHurdles},
and \qref{Q_Maintainability}. Peer review can potentially find misunderstandings
in how the code implements the required theory, which will improve the
software's correctness (\ppref{P_Correctness}). The benefits of peer review for
addressing pain points can be increased by extending the review from just code,
to also reviewing all software artifacts, including documentation, build
scripts, test cases and the development process itself.

@@

We recommend that MI developers use the principle of design for change for their
modularizations.  Although the advice to modularize research software to handle
complexity is common \cite{WilsonEtAl2014, StewartEtAl2017, Storer2017},
specific guidelines on how to divide the software into modules is less
prevalent.  Not every decomposition is a good design for supporting change, as
shown by \cite{Parnas1972a}.  For instance, a design with low cohesion and high
coupling \cite[p.\ 48]{GhezziEtAl2003} will make change difficult. Especially in
research software, where change is inevitable, designers need to produce a
modularization that supports change.  Ocean modelling software is currently
feeling the pain of not emphasizing modularization in legacy code
\cite{JungEtAl2022}.

Developers can accomplish design for change by first identifying likely changes,
either implicitly or explicitly, and second by hiding each likely change behind
a well-defined module interface \cite{Parnas1972a}.

@@

A study is reliable if repeating it by another researcher, using the same
methodology, would lead to the same results~\cite{RunesonAndHost2009}.
We identify the following threats:

\begin{itemize}
\item One individual does the measures for all packages. A different
evaluator might find different results, due to differences in abilities,
experience, and biases.
\item The measurements for the full set of packages took several months (of
elapsed time).  Over this time the software repositories may have changed and
the reviewer's judgement may have drifted.
\end{itemize}

The measurement process used has previously been shown to be reasonably
reproducible.  [citation redacted] 
%% DOUBLE BLIND
%Smith et al.~\cite{SmithEtAl2016} 
reports grading five software
products by two reviewers. Their rankings were almost identical. As long as each
grader uses consistent definitions, the relative comparisons in the AHP results
will be consistent between graders.

@@

Construct validity is when the adopted metrics represent what they are intended
to measure~\cite{RunesonAndHost2009}. We have identified the following
potential issues:
\begin{itemize}
\item We make indirect measurement of software qualities since meaningful direct
measures for qualities like maintainability, reusability and verifiability, are
unavailable.  We follow the usual assumption that developers achieve higher
quality by following procedures and adhering to standards \cite[p.\
112]{VanVliet2000}.
\item We could not install or build \textit{dwv}, \textit{GATE}, and
\textit{DICOM Viewer}. We used a deployed on-line version for \textit{dwv}, a
virtual machine version for \textit{GATE}, but no alternative for \textit{DICOM
Viewer}.
\item Robustness measurements involve only two pieces of data, leading to
limited variation in the robustness scores. Our measurement-time budget limited
what we could achieve here.
\item Our maintainability proxies (higher ratio of comments to source,
high percentage of closed issues) have not been validated.
\item While smaller modules tend to be easier to reuse, small modules are
not necessarily good modules, nor understandable modules.
\item The understandability measure relies on 10 random source code files, but
the 10 files will not necessarily be representative. 
\item Our overall AHP ranking makes the unrealistic assumption of equal
weighting.
\item We approximated popularity by stars and watches.
\item Table~\ref{Tbl_Guidelines} required judgement as
not all guidelines use the same names for artifacts that contain
essentially the same information.
\end{itemize}

@@

While we did find a reasonable amount of documentation, especially when consider
all MI projects, there were definitely some holes. Some important documentation
(test plans and requirements documentation) was missing, and other
(contributors' guide, code of conduct, code style guidelines, product roadmap,
design documentation, and API documentation)
(Section~\ref{Sec_CompareArtifacts}) were rare.

Our interviewees proposed strategies to improve the state of the practice, to
address the identified pain points, and to improve software quality. To their
list (Section~\ref{painpoints}) we added some of our own recommended strategies
(Section~\ref{ch_recommendations}). The strategies that emerged include
increasing documentation, increasing testing by enriching datasets, increasing
modularity, using continuous integration, moving to web applications, employing
linters, increasing peer review, and using design for change to guide
modularization.

A deeper understanding of the needs of the MI community will require data beyond
what is available in repositories.  Future work should involve interviewing MI
developers to better understand their ``pain points''.

>>>>

At least to some extent the problems for MI software users are due to holes in
their knowledge of software and computing technology, and sometimes also
lack expertize in their domain to be able to adequately use the software
\cite{WieseEtAl2019}. 
%A similar pattern has been observed during interviews with
%LBM software developers [citation redacted for blind review].
%\cite{SmithEtAl2024}.

@@

\begin{itemize}
    \item Use documentation (user manuals, mailing lists, forums) 
    \item Usability tests and interviews with end users; and, 
    \item Adjusting the software according to user feedback.
\end{itemize}

@@

\section{Software Qualities}\label{sec:sq}

The structured interviews also included an active discussion of software
qualities, and in particular maintainability and reproducibility. Note that
these did not come up as ``pain points''.
%\cite{SmithEtAl2021}.  
The developers discussed their approaches for improving these qualities, which
we now relay.

\begin{enumerate} \item[Q\refstepcounter{qnum}\theqnum
\label{Q_Maintainability}:] \textbf{Maintainability:}  Maintainability has been
rated as the third most important software quality for research software
\cite{Nguyen-HoanEtAl2010}, and developers complain about 
accumulating too much technical debt \cite{KruchtenEtAl2012}.

The main strategy is to reduce code duplication by using properly structured
libraries.  Other strategies mentioned include supporting third-party
extensions, an easy-to-understand architecture, a dedicated architect, starting
from simple solutions, and documentation.

\item[Q\refstepcounter{qnum}\theqnum \label{Q_Reproducibility}:]
\textbf{Reproducibility:}  
An important precondition for reproducibity is increased and better
documentation (which also helps with many of the afore mentioned pain points).
The challenges of inadequate documentation are a known problem for research
software \cite{PintoEtAl2018, WieseEtAl2019} and for non-research software
\cite{LethbridgeEtAl2003} alike.

The threats to reproducibility that were mentioned include closed-source
software, no user interaction tests, no unit tests, the need to change versions
of some common libraries, variability between CPUs, and misinterpretation of
how manufacturers create medical images.

The most common strategy proposed was
testing (regression tests, unit tests, having good tests). The second most
common strategy is making code, data, and documentation
available, possibly by creating open-source libraries.  Other ideas that were
mentioned include running the same tests on all platforms, a dockerized version
of the software to insulate it from the OS environment, using standard
libraries, monitoring the upgrades of the library dependencies, clearly
documenting the version information, bringing along the exact versions of all
the dependencies with the software, providing checksums of the data, and
benchmarking the software against other software that overlaps in functionality.
Specifically one interviewee suggested using \textit{3D Slicer} as the benchmark
to test their reproducibility.

@@

We provide specific recommendations for future consideration -- these are
not criticisms of past and/or current development practices. We
expand on some of the ideas from our interviews, and bring in further ideas
that could have a noticable impact. The ``new'' ideas are not novel to this
paper but rather from the software engineering and software carpentry domains
that have yet to have seen wide adoption.  We list the ideas roughly in the
order of increasing implementation effort.

@@

Continuous integration involves building and testing the software every
time a coherent change is made to the code (i.e. via a Pull Request)%
\cite[p.\ 13] {HumbleAndFarley2010}, \cite{ShahinEtAl2017, Fowler2006}. Initial
setup can be cumbersome, but
the benefits are impressive:
\begin{itemize}
    \item Elimination of headaches associated with a separate integration phase
    \cite{Fowler2006}, \cite[p.\ 20]{HumbleAndFarley2010}.
    \item Detection and removal of bugs \cite{Fowler2006} via
    automated testing.  
    \item Everyone is always working on a stable base that always passes all
    the tests.
    \item If the CI system uses generators and linters, it will also have
    current documentation and clean code.
\end{itemize}

CI helps relieve several point points:
\begin{enumerate*}
\item by eliminating the time-consuming integration phase, 
freeing more time for development (\ppref{P_LackDevTime}),
\item Automated tests help with ensuring
correctness (\ppref{P_Correctness}) and reproducibility
(\qref{Q_Reproducibility}). 
\end{enumerate*}

\subsection{Perhaps Move To Web Applications} \label{sec_webapps}

Multiple pain points encountered, especially technology hurdles
(\ppref{P_TechnologyHurdles}), could be alleviated by moving to a web
application. But this must be an active decision as it may not be a good
fit in all cases.  The decision needs to be based on whether, on balance, this
would improve the four factors identified earlier: compatibility,
maintainability, performance, and security. We outline the factors to consider
to enable better decision making, as a final decision crucially depends on
each teams' objectives.

@@

\item \textbf{Modern technologies may improve frontend performance.} 
Complex, graphics-heavy user-interfaces often do not respond as quickly
as native applications.  However, new technologies may help. For example,
some JavaScript libraries can help the frontend harness the power of the
computer's GPU and accelerate graphical computing. In addition, there are new
frameworks helping developers with cross-platform compatibility. For example,
the \href{https://flutter.dev/}{Flutter} project enables support for web,
mobile, and desktop OS with one codebase.  Other options include
\href{https://vuejs.org/} {Vue}, \href{https://angular.io/} {Angular} and
\href{https://reactjs.org/} {React}, and \href{https://elm-lang.org/}{Elm}.

\item \textbf{Backend servers can potentially deliver high performance.} If
the principal bottleneck is computational, backend servers may outperform
native applications, as long as latency is not an issue.
Options include
\href{https://www.django-rest-framework.org/} {Django}, \href{https://laravel.com/} {Laravel} and
\href{https://nodejs.org/en/} {Node.js}.
When traffic and latency are an issue,
\href{https://github.com/gin-gonic/gin} {Gin} is an option.

\item \textbf{Some backend servers can be inexpensive.} Serverless solutions
from cloud service providers (like Amazon Web Services (AWS) and Google Cloud
Platform) may be worth exploring. These still use a server, but only actual use
is charged.

\item \textbf{Web transmission may diminish security.} Transferring sensitive
data on-line can be a problem for projects requiring high security, even
if using encryption. Some jurisdiction may not allow transmission of some
health data at all.

@@

Access to real-world imaging datasets can be a testing bottleneck,
jeopardising correctness. Building on developers' suggestions:

\begin{itemize}
\item \textbf{Build and maintain good connections to datasets.} 
Teams can build
connections with medical professionals, who may have
access to private datasets and can perform tests for the team.

\item \textbf{Collect and maintain datasets over time.} It is especially
important to have access to ``unique'' inputs that were outliers in
historical testing scenarios.

\item \textbf{Search for open data sources.} There are many open MI
datasets:
\href{https://nihcc.app.box.com/v/ChestXray-NIHCC}{Chest X-ray Datasets} by
National Institute of Health \cite{WangEtAl2017},
\href{https://www.cancerimagingarchive.net/}{Cancer Imaging Archive}
\cite{PriorEtAl2017}, \href{https://medpix.nlm.nih.gov/home}{MedPix} by
National Library of Medicine \cite{Smirniotopoulos2014}, and datasets for liver
\cite{BilicEtAl2019} and brain \cite{MenzeEtAl2015} tumor segmentation
benchmarks.

\item \textbf{Create sample data for testing.} If a team is able to create
baseline and/or synthetic data sets with known characteristics, this is a
worthwhile investment (and should hopefully be made public). For example, using
an MRI scanner to create images of objects, animals, and volunteers who waive
the privacy rights to their image (i.e. not patients).

%JC: I'm aware of too many caveats wrt this recommendation, so I am taking
% it out.
%
%\item \textbf{Remove privacy from sensitive data.} For data with sensitive
%information, a team can ask the data owner to remove such information or add
%noise to protect privacy. One example is using de-identified copies of medical
%images for testing.

@@

\subsection{Use Linters and other Static Analysis Tools} \label{Sec_Linters}

Linters are simple tools that analyzes code to find programming errors,
suspicious constructs, and stylistic inconsistencies \cite{Wikipedia2022_Lint}.
While they can be used in an ad hoc fashion, coupling them with CI is where
the largest gain happens. Automatic use of code quality tools is increasing
commonplace for industrial software -- it is regretable that the
research software guidlines have no caught up to this.

These tools have many benefits: finding potential bugs, finding memory
leaks, improving performance, standardizing code with respect to formatting,
removing silly errors before code reviews, and catching potential security
issues~\cite{SourceLevel2022_Lint}. Most popular programming languages possess
such tools. This helps with \ppref{P_LackDevTime}, \qref{Q_Maintainability},
and correctness (\ppref{P_Correctness}).

The pervasive use of such tools to enforce documentation standards
partially explains the relatively higher quality of statistical tools that are
part of the Comprehensive R Archive Network (CRAN) [citation redacted for blind
review].  %\cite{SmithEtAl2018_StatSoft}.

\subsection{Conduct Peer Reviews} \label{Sec_PeerReview}

Peer review is recommended by many \cite{HerouxEtAl2008,
OrvizEtAl2017, USGS2019}. Modern peer review processes are lightweight,
informal, tool-based, asynchronous, and focused
on reviewing code changes~\cite{SadowskiEtAl2018}. GitHub, for example,
provides such tools for its pull requests.

Rigorous inspection finds 60-65\% of latent defects on
average, and often tops 85\% in defect removal efficiency \cite{Jones2008}. The
success rate of code inspection is generally higher than most forms of testing,
which average between 30 --- 35\% for defect removal efficiency
\cite{EbertAndJones2009, Jones2008}. For research software, Kelly and Shepard
\cite{KellyAndShepard2000} show a task based inspection approach can be
effective.

Due to improving code quality and increasing knowledge transfer, peer review
addresses the same pain points and qualities as linters \ppref{P_LackDevTime},
\qref{Q_Maintainability} and \ppref{P_Correctness}. The benefits can be
increased by extending reviews to all artifacts, including documentation, build
scripts, test cases and the development process itself.

@@

 Although the advice to modularize research software to handle
complexity is common \cite{WilsonEtAl2014, StewartEtAl2017, Storer2017},
specific guidelines on how to divide the software into modules is less
prevalent.  Not every decomposition is a good design for supporting change, as
shown by \cite{Parnas1972a}.  For instance, a design with low cohesion and high
coupling \cite[p.\ 48]{GhezziEtAl2003} will make change difficult. Especially in
research software, where change is inevitable, designers need to produce a
modularization that supports change.  Ocean modelling software is currently
feeling the pain of not emphasizing modularization in legacy code
\cite{JungEtAl2022}.

@@

A study is reliable if repeating it by another researcher, using the same
methodology, would lead to the same results~\cite{RunesonAndHost2009}.
We identify the following threats:
\begin{enumerate*}
\item One individual does the measures for all packages. A different
evaluator might find different results, due to differences in abilities,
experience, and biases.
\item The measurements for the full set of packages took several months (of
elapsed time).  Over this time the software repositories may have changed and
the reviewer's judgement may have drifted.
\end{enumerate*}

The measurement process used has previously been shown to be reasonably
reproducible, [citation redacted]
%% DOUBLE BLIND
%Smith et al.~\cite{SmithEtAl2016} 
reports grading five software products by two reviewers. Their rankings were
almost identical. As long as each grader uses consistent definitions, the
relative comparisons in the AHP results will be consistent between graders.

@@

Construct validity is when the adopted metrics represent what they are intended
to measure~\cite{RunesonAndHost2009}. We have identified the following
potential issues:
\begin{itemize}
\item We make indirect measurement of software qualities since meaningful direct
measures for qualities like maintainability, reusability and verifiability, are
unavailable.  We follow the usual assumption that developers achieve higher
quality by following procedures and adhering to standards \cite[p.\
112]{VanVliet2000}.
\item We could not do all measurements for some software as we could not
install or build \textit{dwv}, \textit{GATE}, and \textit{DICOM Viewer}. We
used a deployed on-line version for \textit{dwv}, a virtual machine version for
\textit{GATE}, but no alternative for \textit{DICOM Viewer}.
\item Robustness measurements involve only two pieces of data, leading to
limited variation in the robustness scores. Our measurement-time budget limited
what we could achieve here.
\item Our maintainability proxies (higher ratio of comments to source,
high percentage of closed issues) have not been validated.
\item While smaller modules tend to be easier to reuse, small modules are
not necessarily good modules, nor understandable modules.
\item The understandability measure relies on 10 random source code files, but
the 10 files will not necessarily be representative. 
\item Our overall AHP ranking makes the unrealistic assumption of equal
weighting.
\item We approximated popularity by stars and watches.
\item Table~\ref{Tbl_Guidelines} required judgement as
not all guidelines use the same names for artifacts that contain
essentially the same information.
\end{itemize}

@@

While we did find a reasonable amount of documentation, especially when consider
all MI projects, there were definitely some holes. Some important documentation
(test plans and requirements documentation) was missing, and other
(contributors' guide, code of conduct, code style guidelines, product roadmap,
design documentation, and API documentation) were rare.

Our interviewees proposed strategies to improve the state of the practice, to
address the identified pain points, and to improve software quality. To their
list (Section~\ref{painpoints}) we added some of our own recommended strategies
(Section~\ref{ch_recommendations}). The strategies that emerged include
increasing documentation, increasing testing by enriching datasets, increasing
modularity, using continuous integration, moving to web applications, using
linters, increasing peer review, and using design for change to guide
modularization.

A deeper understanding of the needs of the MI community will require data beyond
what is available in repositories.
